---
title: "Glider Deployment Report: AMLR01 (November 20, 2024)"
format: 
  pdf: 
    toc: true
    colorlinks: true
    papersize: letter
---

### Summary

```{r, warning = FALSE}
#| echo: false
#| message: false
#| include: false
library(ncdf4)
library(dplyr)
library(lubridate)
library(here)
library(swfscMisc)
library(cowplot)
library(googleway)
library(ggplot2)
library(ggrepel)
library(ggspatial)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(lwgeom)
library(knitr)
library(stringr)
library(tidyr)
library(readr)
library(kableExtra)

#enter deployment name nere, comment in deployment location
deployment<-"amlr01-20241120"
#locale <- "CCE"
locale <- "Antarctica"

#comment in/out battery capacity as appropriate
#std.pri<-550
#std.rec<-215
ext.pri<-800
#ext.rec<-300

#comment in/out sensors as appropriate
ctd.file<-"sample01.ma"
puck.file<-"sample48.ma"
oxy.file<-"sample54.ma"
azfp.file<-"sample68.ma"
#nortek.file<-"sample64.ma"
#sg.file<-"sample87.ma"
#wispr.file<-"sample93.ma"
#dmon.file<-"sample49.ma"
#par.file<-"sample56.ma"
ctd.spec<-"Sea-Bird Conductivity Temperature Depth (CTD) (SBE-41)"
puck.spec<-"Sea-Bird ECO Puck (backscatter and fluorescence) (FLBBCD-SLC, CDOM)"
oxy.spec<-"AANDERAA oxygen optode (4831)"
azfp.spec<-"Acoustic Zooplankton Fish Profiler (ASL)"
#nortek.spec<-"Mini Signature-100 wideband echosounder (Nortek)"
#sg.spec<-"Shadowgraph camera (Williamson and Associates)"
#wispr.spec<-"Passive acoustic monitor (Wispr3, Embedded Ocean Systems)"
#dmon.spec<-"Passive acoustic monitor (DMON2, Woods Hole Oceanographic Institution)"
#par.spec<-"Photosynthetically active radiation (QSP-2150, Biospherical Instruments, Inc.)"
state.to.sample.7<-"Diving, hovering, climbing"
#state.to.sample.15<-"Diving, hovering, climbing, on surface"
sample.depth<-"1000 m"
```

```{r}
#| echo: false
#| message: false
#| include: false

#change file names
#set base path, then use Google Cloud Shell to download files into folders within base path
base.path <- "deployment-reports/Data"
#update as necessary. For sensors with .cfg files, comment in the line(s) associated with the sensor
#listpacfg = wispr PAM
#listaacfg = active acoustics (Nortek or AZFP)
#listsgcfg = shadowgraph camera
eng.ts<-here(base.path,deployment,"processed-L1",paste0(deployment,"-delayed-eng.nc"))
raw.ts<-here(base.path,deployment,"processed-L1",paste0(deployment,"-delayed-raw.nc"))
listma<-list.files(here(base.path,deployment,"archive-sfmc"),pattern="*_sample*")
listpacfg<-list.files(here(base.path,deployment,"archive-sfmc"),pattern="*_wispr.cfg")
listaacfg<-list.files(here(base.path,deployment,"archive-sfmc"),pattern="*_ad2cp.cfg")
listaacfg<-list.files(here(base.path,deployment,"archive-sfmc"),pattern="*_azfp.cfg")
listsgcfg<-list.files(here(base.path,deployment,"archive-sfmc"),pattern="*_solocam.cfg")
nc.eng <- nc_open(eng.ts)
nc.raw <- nc_open(raw.ts)

nc.eng.time <- as.POSIXct(ncvar_get(nc.eng, "time"), origin = "1970-01-01")
nc.eng.lat <- ncvar_get(nc.eng, "latitude")
nc.eng.lon <- ncvar_get(nc.eng, "longitude")
nc.raw.lat <- ncvar_get(nc.raw, "gps_latitude")
nc.raw.lon <- ncvar_get(nc.raw, "gps_longitude")

lat<-nc.raw.lat
lon<-nc.raw.lon
surf<-data.frame(
  lat=nc.raw.lat,
  lon=nc.raw.lon
)
surf<-na.omit(surf)

surf <- surf %>% 
  filter(between(lat, -90, 90), between(lon, -180, 180)) %>% 
  mutate(lat2 = lead(lat, 1), 
         lon2 = lead(lon, 1)) %>% 
  slice_head(n = -1)

distances<-swfscMisc::distance(surf$lat,surf$lon,surf$lat2,surf$lon2,
                               method="haversine", units = "km")
 

#sum for total distance covered
sum(distances)

# code that splits date from time to calculate days deployed
glider.date<-as.character(nc.eng.time)
new.glider.date<-strsplit(glider.date,split=" ")
new.glider.date<-unlist(new.glider.date)
new.glider.date<-new.glider.date[c(TRUE,FALSE)]
deployment.length<-unique(new.glider.date)
days.deployed<-length(deployment.length)

```

The Ecosystem Science Division (ESD) at the Southwest Fisheries Science Center (SWFSC) deployed glider **AMLR01** (unit_681) on `{r} deployment.length[1]` off the coast of **Smith Island in the Southern Ocean** (`{r} round(nc.eng.lat[1],digits=2)`^o^N, `{r} round(nc.eng.lon[1],digits=2)`^o^W) (Figure 1). Sensors deployed on the glider are listed in @tbl-sensors.

```{r}
#| echo: false
#| message: false

#code for number of dives performed during deployment
nc.eng.yos <- ncvar_get(nc.eng, "total_num_inflections")
nc.eng.yos<-as.vector(nc.eng.yos)
nc.eng.yos<-na.omit(nc.eng.yos)
dives<-round((max(nc.eng.yos)-min(nc.eng.yos))/2,digits=0)

#max depth
nc.eng.depth <- ncvar_get(nc.eng, "depth")
nc.eng.depth <- as.vector(nc.eng.depth)
max.depth <- max(nc.eng.depth)
```

@fig-map-1 (A) displays tracklines of glider during deployment, while (B) displays the broad deployment area. The glider remained deployed for `{r} days.deployed` days, performed `{r} dives` dives, and traveled a total of `{r} round(sum(distances),digits=2)` km while diving to a maximum depth of `{r} round(max.depth,digits=1)` m. The glider was recovered on `{r} deployment.length[length(deployment.length)]`, approximately 25 km southwest of Nelson Island.

```{r, warning = FALSE}
#| label: fig-map-1
#| fig-cap: "Glider tracklines. A displays close-up tracklines, while B displays the broad deployment area."
#| fig-alt: "The map on the left shows a close-up display of where the glider surfaced throughout its deployment. The map on the right is zoomed out to show the approximate location of the glider deployment relative to land."
#| echo: false
#| message: false
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 

world <- ne_countries(scale = "medium", returnclass = "sf")

minlat<-min(surf$lat)-0.2
maxlat<-max(surf$lat)+0.2
minlon<-min(surf$lon)-0.5
maxlon<-max(surf$lon)+0.5
p1<-ggplot(data = world) +
  geom_sf() +
  geom_point(data = surf, aes(x = lon, y = lat), color = "darkred", size=1) +
  coord_sf(xlim = c(maxlon,minlon), ylim = c(maxlat, minlat), expand = FALSE)+
  theme(axis.text.x=element_text(angle=90))+
  theme(panel.background=element_rect(fill=NA),panel.grid.major=element_line(color="grey50"))+
  xlab("Longitude")+ylab("Latitude")

if (locale == "CCE") {
  minlon2 <- -126
  maxlon2 <- -115
  minlat2 <- 30
  maxlat2 <- 45
} else if (locale == "Antarctica") {
  minlon2 <- -80
  maxlon2 <- -50
  minlat2 <- -66
  maxlat2 <- -50
} else {
  stop("invalid value for 'locale'")
}

p2<-ggplot(data = world) +
  geom_sf() +
  geom_point(data = surf, aes(x = lon, y = lat), color = "darkred", size=1) +
  coord_sf(xlim = c(maxlon2,minlon2), ylim = c(maxlat2, minlat2), expand = FALSE)+
  theme(axis.text.x=element_text(angle=90))+
  theme(panel.background=element_rect(fill=NA),panel.grid.major=element_line(color="grey50"))+
  xlab("Longitude")+ylab("Latitude")

plot_grid(p1, p2, labels = c('A','B'))
```

```{r}
#| label: tbl-sensors
#| tbl-cap: "Science sampling strategies for current glider deployment. Additional settings for the AZFP echosounder, the Williamson and Associates camera, and the Nortek echosounder (if installed) are defined in configuration and initialization files on the glider's science computer, and are also housed on the Google Cloud Platform. All deployment files are available on request."
#| echo: false
#| message: false

sensors<-data.frame()
sensors %>%
  reframe(
    `File Name` = c(ctd.file,puck.file,oxy.file,azfp.file),
    `Sensor` = c(ctd.spec,puck.spec,oxy.spec,azfp.spec),
    `State to Sample` = rep(state.to.sample.7,4),
    #`State to Sample` = c(rep(state.to.sample.7,3),state.to.sample.15),
    `Depth to Sample` = rep(sample.depth,4),
    `Serial Number` = c("9547","5583","0643","59010")
    ) %>%
  knitr::kable(format="latex",align="l",booktabs=TRUE,linesep="") %>%
  kableExtra::kable_styling(position="left",latex_options="striped",stripe_color="gray!15") %>%
  column_spec(1,width="1in") %>%
  column_spec(2,width="2in") %>%
  column_spec(3,width="1in") %>%
  column_spec(4,width="1in") %>%
  column_spec(5,width="1in")
```

This deployment had two goals: 1) to estimate the density of Antarctic krill in an area important for krill-dependent predators and the krill fishery; and 2) to compare echosounder data between the older Acoustic Zooplankton Fish Profiler (AZFP, deployed on AMLR01) and the newer Nortek wideband echosounder (concurrently deployed on AMLR08). AMLR08 suffered a pump malfunction and was recovered on 2024-11-24.

### Introduction

The Ecosystem Science Division at NOAA Fisheries’ Southwest Fisheries Science Center monitors the living marine resources within the Southern Ocean and the California Current in order to satisfy the requirements of several legislative mandates to support conservation and management decision-making. To achieve this goal, we use autonomous underwater buoyancy-driven gliders with integrated sensors for measuring ocean conditions, plankton densities, and marine mammal distributions.

Depending on the specific deployment objective, Slocum gliders are equipped with a suite of sensors. We obtain acoustic estimates of zooplankton density (primarily Antarctic krill in the Southern Ocean) using one of two different echosounders: an Acoustic Zooplankton Fish Profiler with discrete frequencies at 67.5 and 125 kHz (AZFP, ASL, Inc) and a mini-Signature 100 wideband echosounder with continuous frequencies between 70 and 120 kHz (Nortek). We also collect ancillary oceanographic data (temperature, salinity, dissolved oxygen, chlorophyll, colored dissolved organic matter, backscatter, and photosynthetically active radiation) to characterize the marine environment. Additional sensors may include passive acoustic monitors for marine mammal detection (“Wispr”, Embedded Ocean Systems; digital acoustic monitoring “DMON”, Woods Hole Oceanographic Institution), “glidercams” for verifying acoustic targets (Williamson and Associates, Inc.) and shadowgraph cameras for obtaining imagery of the plankton community (Williamson and Associates, Inc.). Imagery is used to train artificial intelligence (AI) models to automate plankton identification.


### Pre-deployment preparation and testing

Prior to deployment, the ESD has a standard protocol for preparing and testing gliders to minimize or eliminate issues that may occur due to human error during deployment:

*Slocum gliders*

1.  Gliders are properly ballasted (i.e., weighted) so that the density of the glider matches the density of the water in which it will be deployed. Weight and flotation configurations are documented
2.  The junctions between glider sections are thoroughly cleaned, old o-rings are discarded, new o-rings are inspected for damage that may compromise their ability to form a water-tight seal, new o-rings are properly lubricated, and the glider is sealed together. All cable connections are photographed to document the “final seal” and ensure the glider was reassembled properly
3.  A “Functional Checkout” is performed to ensure and document that all glider systems and science sensors are functioning properly. During the Functional Checkout, we verify the battery type installed in the glider (lithium primary or lithium rechargeable) and that the appropriate battery duration (total coulomb amp hours) is active in the glider’s autoexec.mi file
4.  Two test missions are performed in the SWFSC test tank (20 m x 10 m x 10 m) to ensure the glider is performing as expected
5.  Once per year, glider compasses are calibrated (the compass was not calibrated prior to this deployment)
6.  Biofouling prevention measures are applied as necessary
7.  Glider flight and science sampling files are prepared according to mission objectives. These objectives are identified by the Principal Investigators for each deployment. Files are uploaded to the Teledyne Webb Research Slocum Fleet Mission Control (SFMC) web interface and sent to the glider just prior to deployment over the Iridium connection
8.  When gliders are shipped to their deployment location, ESD glider technicians perform a second Functional Checkout to ensure the gliders function properly after transit


### Deployment-specific testing

Because AMLR01 had not been deployed since 2021 and AMLR08 had not been deployed since it was repaired after a catastrophic electrical failure in 2022, both gliders were deployed in San Diego for one day in July to ensure they performed as expected. Both gliders performed well, and because their sensor configurations did not include any combinations of sensors that had led to electrical problems in the past, no further deployment-specific testing (beyond our standard pre-deployment preparations and tests) was done.

### Deployment

“AMLR01” was deployed on `{r} deployment.length[1]`, off the coast of Smith Island in the Southern Ocean (`{r} round(nc.eng.lat[1],digits=2)`^o^N, `{r} round(nc.eng.lon[1],digits=2)`^o^W) from the DAP vessel M/V *Betanzos*. This glider was deployed with the sensor configuration listed above, and with lithium primary batteries (coulomb amp hour total = 800). We began this deployment using the autoballast feature to maximize oil pump efficiency while maintaining the appropriate dive angle of 23 degrees for the AZFP echosounder. Autoballast converged successfully for the majority of the deployment. 

```{r}
#| echo: false
#| message: false

#for battery usage. Must fill in proper coulomb amphr total for ca.percent.used
nc.eng.cat<-ncvar_get(nc.eng,"total_amphr")
nc.eng.cat<-as.vector(nc.eng.cat)
ca.used<-max(nc.eng.cat)-min(nc.eng.cat)
ca.percent.used<-(ca.used/ext.pri)*100
```

This glider performed well for the entire `{r} days.deployed`-day mission. The glider used `{r} round(ca.used,digits=2)` amp hours over `{r} days.deployed` days, or `{r} round(ca.percent.used,digits=2)`% of its battery capacity.

This was the U.S. AMLR Program’s first glider deployment with no mission interruptions (i.e., aborts). While the glider experienced strong eastward currents along the southern coast of King George Island, pilots were able to direct the glider into deeper water and keep it on track. The glider was recovered on schedule on `{r} deployment.length[length(deployment.length)]`, approximately 25 km southwest of Nelson Island, by the M/Y *Hanse Explorer* and U.S. AMLR Program staff. 

The only noteworthy “issue” of this deployment occurred when a military glider with a different name experienced an abort and connected to Teledyne Webb’s Slocum Fleet Mission Control as AMLR01, briefly leading us to believe that our glider had aborted its mission. We were able to quickly determine that our AMLR01 had not aborted its mission.

AMLR01 remains in Antarctica at Cape Shirreff until camp closes in March, at which point it will be transported to Punta Arenas, Chile, and eventually back to the U.S.


### Post-deployment actions

We received AMLR01 back in La Jolla, CA, on 2025-06-12. The glider was inspected, cleaned, and all data were downloaded and archived on NOAA's Google Cloud Platform. Data will be quality checked and analyzed at a later date.

```{r}
bs<-here(base.path,deployment,"plots","delayed","spatialGrids-sci","amlr01-20241120_backscatter_700_spatialGrids.png")
chl<-here(base.path,deployment,"plots","delayed","spatialGrids-sci","amlr01-20241120_chlorophyll_spatialGrids.png")
den<-here(base.path,deployment,"plots","delayed","spatialGrids-sci","amlr01-20241120_density_spatialGrids.png")
oxy<-here(base.path,deployment,"plots","delayed","spatialGrids-sci","amlr01-20241120_oxygen_concentration_spatialGrids.png")
sal<-here(base.path,deployment,"plots","delayed","spatialGrids-sci","amlr01-20241120_salinity_spatialGrids.png")
temp<-here(base.path,deployment,"plots","delayed","spatialGrids-sci","amlr01-20241120_temperature_spatialGrids.png")
```

### Figures

Plots below are generated from raw data which has not yet been quality-checked. 

![a](`{r} bs`){fig-alt="A plot showing points where the glider surfaced by latitude and longitude, with two color-coded plots of backscatter by latitude and depth and by longitude and depth."} 
![b](`{r} chl`){fig-alt="A plot showing points where the glider surfaced by latitude and longitude, with two color-coded plots of chlorophyll-a by latitude and depth and by longitude and depth."} 
![c](`{r} den`){fig-alt="A plot showing points where the glider surfaced by latitude and longitude, with two color-coded plots of water density by latitude and depth and by longitude and depth."} 
![d](`{r} oxy`){fig-alt="A plot showing points where the glider surfaced by latitude and longitude, with two color-coded plots of oxygen concentration by latitude and depth and by longitude and depth."} 
![e](`{r} sal`){fig-alt="A plot showing points where the glider surfaced by latitude and longitude, with two color-coded plots of salinity by latitude and depth and by longitude and depth."} 
![f](`{r} temp`){fig-alt="A plot showing points where the glider surfaced by latitude and longitude, with two color-coded plots of temperature by latitude and depth and by longitude and depth."}


