---
title: "Glider Deployment Report: unit_1024 (March 9th, 2026)"
format: 
  pdf: 
    toc: true
    colorlinks: true
    papersize: letter
---

```{r, warning = FALSE}
#| label: setup 
#| echo: false 
#| message: false
#| include: false

# load packages
library(ncdf4)
library(dplyr)
library(lubridate)
library(here)
library(swfscMisc)
library(cowplot)
library(googleway)
library(ggplot2)
library(ggrepel)
library(ggspatial)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(lwgeom)
library(knitr)
library(stringr)
library(tidyr)
library(readr)
library(kableExtra)

# edit set global variables
glider_name <- "unit_1024" 
glider_serial <- "1024" 
deployment <-"unit_1024-20260309"
deployment_year <- "2026"
deployment_date <- "March 9, 2026"
locale <- "CCE" # "CCE", "Antarctica", "Hawaii"

battery_type <- "lithium rechargeable extended" # "lithium primary", "lithium primary extended", "lithium rechargeable", "lithium rechargeable extended"
battery_capacity <- "300" # "720", "1050", "215", "300"

#comment in/out sensors as appropriate and add serial numbers
ctd.file <- "sample01.ma"
puck.file <- "sample48.ma"
oxy.file <- "sample54.ma"
#azfp.file <- "sample68.ma"
#nortek.file <- "sample64.ma"
sg.file <- "sample87.ma"
#wispr.file <- "sample93.ma"
#dmon.file <- "sample49.ma"
#par.file <- "sample56.ma"

ctd.spec <- "Sea-Bird Conductivity Temperature Depth (CTD) (SBE-41)"
puck.spec <- "Sea-Bird ECO Puck (backscatter and fluorescence) (FLBBCD-SLC, CDOM)"
oxy.spec <- "AANDERAA oxygen optode (4831)"
#azfp.spec <- "Acoustic Zooplankton Fish Profiler (ASL)"
#nortek.spec <- "Mini Signature-100 wideband echosounder (Nortek)"
sg.spec <- "Shadowgraph camera (Williamson and Associates)"
#wispr.spec <- "Passive acoustic monitor (Wispr3, Embedded Ocean Systems)"
#dmon.spec <- "Passive acoustic monitor (DMON2, Woods Hole Oceanographic Institution)"
#par.spec <- "Photosynthetically active radiation (QSP-2150, Biospherical Instruments, Inc.)"

ctd.sn    <- "9715"
puck.sn   <- "6866"
oxy.sn    <- "1160"
#azfp.sn  <- ""
#nortek.sn<- ""
sg.sn    <- "SG-V3-003"
#wispr.sn <- ""
#dmon.sn   <- ""
#par.sn    <- ""

state.to.sample.7 <- "Diving, hovering, climbing"
state.to.sample.15 <- "Diving, hovering, climbing, on surface"
sample.depth <- "1000 m"
sample.depth.sg <- "300 m"

# set where script is being run, GCP workstation or local machine
processor <- "Local" # "GCP", "Local"
```

```{r}
#| label: access data files
#| echo: false
#| message: false
#| include: false

if (processor == "GCP") {
  # workstation - access GCP bucket files
  home_dir <- Sys.getenv("HOME")
  base.path <- file.path(home_dir, "gcs-mnt", "swfscesd-glider-deployments-data-out")

  eng.ts <- file.path(base.path, deployment_year, deployment, "processed-L1", paste0(deployment,"-delayed-eng.nc"))
  raw.ts <- file.path(base.path, deployment_year, deployment, "processed-L0", paste0(deployment,"-delayed-raw.nc"))
  
  # TODO - ma files, currently in cold storage so still accessing locally
  local.path <- "/home/user/glider-lab/deployment-reports/data"
  listma <- list.files(local.path, pattern="sample.*\\.ma$")
  # listpacfg <- list.files(local.path, pattern="*_wispr.cfg")
  # listaacfg <- list.files(local.path, pattern="*_ad2cp.cfg")
  # listaacfg <- list.files(local.path, pattern="*_azfp.cfg")
  listsgcfg <- list.files(local.path, pattern="*_solocam.cfg")
  
} else if (processor == "Local") {
  #  local - access local files saved to the deployment-reports/Data folder 
  #set base path, then use Google Cloud Shell to download files into folders within base path
  base.path <- "Data"

    eng.ts<-here(base.path,deployment,paste0(deployment_year, "_", deployment, "_processed-L1_", deployment, "-delayed-eng.nc"))
  raw.ts<-here(base.path,deployment,paste0(deployment_year, "_", deployment, "_processed-L0_", deployment, "-delayed-raw.nc"))
  
  # update as necessary. For sensors with .cfg files, comment in the line(s) associated with the sensor

  listma <- list.files(here(base.path, deployment), pattern="sample.*\\.ma$")
  # listpacfg <- list.files(here(base.path, deployment), pattern="*_wispr.cfg")
  # listaacfg <- list.files(here(base.path, deployment), pattern="*_ad2cp.cfg")
  # listaacfg <- list.files(here(base.path, deployment), pattern="*_azfp.cfg")
  listsgcfg <- list.files(here(base.path, deployment), pattern="solocam.cfg")
}

nc.eng <- nc_open(eng.ts)
nc.raw <- nc_open(raw.ts)

nc.eng.time <- as.POSIXct(ncvar_get(nc.eng, "time"), origin = "1970-01-01")
nc.eng.lat <- ncvar_get(nc.eng, "latitude")
nc.eng.lon <- ncvar_get(nc.eng, "longitude")
nc.raw.lat <- ncvar_get(nc.raw, "gps_latitude")
nc.raw.lon <- ncvar_get(nc.raw, "gps_longitude")

lat<-nc.raw.lat
lon<-nc.raw.lon
surf<-data.frame(
  lat=nc.raw.lat,
  lon=nc.raw.lon
)
surf<-na.omit(surf)

#eliminate erroneous gps hits to calculate an accurate distance. May need to visualize on a map and comment these in and out as necessary  
#minlon.thresh<-min(surf$lon)+0.05
#maxlon.thresh<-max(surf$lon)-0.05
#surf<-surf[surf$lon > minlon.thresh & surf$lon < maxlon.thresh,]

#minlat.thresh<-min(surf$lat)+0.05
#maxlat.thresh<-max(surf$lat)-0.05 
#surf<-surf[surf$lat > minlat.thresh & surf$lat < maxlat.thresh,]

#calculating distance traveled from GPS points
#this creates 2 vectors that are offset so that the first row of lat and lon is the starting point and the first row of lat2 and lon2 is the ending point

surf <- surf %>% 
  filter(between(lat, -90, 90), between(lon, -180, 180)) %>% 
  mutate(lat2 = lead(lat, 1), 
         lon2 = lead(lon, 1)) %>% 
  slice_head(n = -1)

distances<-swfscMisc::distance(surf$lat,surf$lon,surf$lat2,surf$lon2,
                               method="haversine", units = "km")
 

#sum for total distance covered
sum(distances)

# code that splits date from time to calculate days deployed
glider.date<-as.character(nc.eng.time)
new.glider.date<-strsplit(glider.date,split=" ")
new.glider.date<-unlist(new.glider.date)
new.glider.date<-new.glider.date[c(TRUE,FALSE)]
deployment.length<-unique(new.glider.date)
days.deployed<-length(deployment.length)

```

## Summary

The Ecosystem Science Division (ESD) at the Southwest Fisheries ScienceCenter (SWFSC) deployed glider `{r} glider_name` (`{r} glider_serial`) on `{r} deployment_date` off the coast of Southern California (`{r} round(nc.eng.lat[1],digits=2)`^o^N, `{r} round(nc.eng.lon[1],digits=2)`^o^W) (Figure 1). Sensors deployed on the glider are listed in @tbl-sensors.

```{r}
#| label: dives
#| echo: false
#| message: false

#code for number of dives performed during deployment
nc.eng.yos <- ncvar_get(nc.eng, "profile_index")
# nc.eng.yos <- ncvar_get(nc.eng, "total_num_inflections") #not a valid variable name, replaced above by 'profile_index'
nc.eng.yos <- as.vector(nc.eng.yos)
nc.eng.yos <- na.omit(nc.eng.yos)
dives <- round((max(nc.eng.yos)-min(nc.eng.yos))/2,digits=0)

#max depth
nc.eng.depth <- ncvar_get(nc.eng, "depth")
nc.eng.depth <- as.vector(nc.eng.depth)
max.depth <- max(nc.eng.depth)
```

@fig-map-1 (A) displays tracklines of glider during deployment, while (B) displays the broad deployment area relative to land. The glider remained deployed for `{r} days.deployed` days, performed `{r} dives` dives, and traveled a total of `{r} round(sum(distances),digits=2)` km while diving to a maximum depth of `{r} round(max.depth,digits=1)` m. No mission aborts occurred and the glider was recovered on `{r} deployment.length[length(deployment.length)]`.

The primary goal of this deployment was to test a new Shadowgraph model (SG-V3-02) to obtain imagery of the plankton community, in preparation for two extended deployments planned to support the upcoming spring California Cooperative Oceanic Fisheries Investigations (CalCOFI) cruise. The shadowgraph camera was programmed to take photos every 2 seconds on dives and climbs, with a 50% red flash brightness. During the deployment, the shadowgraph camera took 30,236 photos and used 3.4% of the camera storage (63.2 GB). No mission aborts occurred and the glider was successfully recovered as scheduled on March 10, 2026.

```{r, warning = FALSE}
#| label: fig-map-1
#| fig-cap: "Glider tracklines. A displays close-up tracklines, while B displays the broad deployment area."
#| fig-alt: "The map on the left shows a close-up display of where the glider surfaced throughout its deployment. The map on the right is zoomed out to show the approximate location of the glider deployment relative to land."
#| echo: false
#| message: false
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 

world <- ne_countries(scale = "medium", returnclass = "sf")

minlat<-min(surf$lat)
maxlat<-max(surf$lat)
minlon<-min(surf$lon)
maxlon<-max(surf$lon)
p1<-ggplot(data = world) +
  geom_sf() +
  geom_point(data = surf, aes(x = lon, y = lat), color = "darkred", size=1) +
  coord_sf(xlim = c(maxlon,minlon), ylim = c(maxlat, minlat), expand = FALSE)+
  theme(axis.text.x=element_text(angle=90))+
  theme(panel.background=element_rect(fill=NA),panel.grid.major=element_line(color="grey50"),panel.ontop=TRUE)+
  xlab("Longitude")+ylab("Latitude")

if (locale == "CCE") {
  minlon2 <- -126
  maxlon2 <- -115
  minlat2 <- 30
  maxlat2 <- 45
} else if (locale == "Antarctica") {
  minlon2 <- -62
  maxlon2 <- -57
  minlat2 <- -63
  maxlat2 <- -62.5
} else {
  stop("invalid value for 'locale'")
}

p2<-ggplot(data = world) +
  geom_sf() +
  geom_point(data = surf, aes(x = lon, y = lat), color = "darkred", size=1) +
  coord_sf(xlim = c(maxlon2,minlon2), ylim = c(maxlat2, minlat2), expand = FALSE)+
  theme(axis.text.x=element_text(angle=90))+
  theme(panel.background=element_rect(fill=NA),panel.grid.major=element_line(color="grey50"))+
  xlab("Longitude")+ylab("Latitude")

plot_grid(p1, p2, labels = c('A','B'))
```

```{r}
#| label: tbl-sensors
#| tbl-cap: "Science sampling strategies for current glider deployment. Additional settings for the AZFP echosounder, the Williamson and Associates camera, and the Nortek echosounder (if installed) are defined in configuration and initialization files on the glider's science computer, and are also housed on the Google Cloud Platform. All deployment files are available on request."
#| echo: false
#| message: false

# build table based on what sensors are active in the setup chunk
sensors <- bind_rows(
  if(exists("ctd.file")) tibble(`File Name`=ctd.file, Sensor=ctd.spec, `State to Sample`=state.to.sample.7, `Depth to Sample`=sample.depth, `Serial Number`=ctd.sn),
  if(exists("puck.file")) tibble(`File Name`=puck.file, Sensor=puck.spec, `State to Sample`=state.to.sample.7, `Depth to Sample`=sample.depth, `Serial Number`=puck.sn),
  if(exists("oxy.file")) tibble(`File Name`=oxy.file, Sensor=oxy.spec, `State to Sample`=state.to.sample.7, `Depth to Sample`=sample.depth, `Serial Number`=oxy.sn),
  if(exists("par.file")) tibble(`File Name`=par.file, Sensor=par.spec, `State to Sample`=state.to.sample.7, `Depth to Sample`=sample.depth, `Serial Number`=par.sn),
  if(exists("dmon.file")) tibble(`File Name`=dmon.file, Sensor=dmon.spec, `State to Sample`=state.to.sample.7, `Depth to Sample`=sample.depth, `Serial Number`=dmon.sn),
  if(exists("sg.file")) tibble(`File Name`=sg.file, Sensor=sg.spec, `State to Sample`=state.to.sample.7, `Depth to Sample`=sample.depth.sg, `Serial Number`=sg.sn),
  if(exists("wispr.file")) tibble(`File Name`=wispr.file, Sensor=wispr.spec, `State to Sample`=state.to.sample.7, `Depth to Sample`=sample.depth, `Serial Number`=wispr.sn),
  if(exists("azfp.file")) tibble(`File Name`=azfp.file, Sensor=azfp.spec, `State to Sample`=state.to.sample.7, `Depth to Sample`=sample.depth, `Serial Number`=azfp.sn),
  if(exists("nortek.file")) tibble(`File Name`=nortek.file, Sensor=nortek.spec, `State to Sample`=state.to.sample.7, `Depth to Sample`=sample.depth, `Serial Number`=nortek.sn)
)

# render table
sensors %>%
  knitr::kable(format="latex", align="l", booktabs=TRUE, linesep="") %>%
  kableExtra::kable_styling(position="left", latex_options="striped", stripe_color="gray!15") %>%
  column_spec(1, width="1in") %>%
  column_spec(2, width="2in") %>%
  column_spec(3, width="1in") %>%
  column_spec(4, width="1in") %>%
  column_spec(5, width="1in")

# sensors<-data.frame()
# sensors %>%
#   reframe(
#     `File Name` = c(ctd.file,puck.file,oxy.file,par.file,dmon.file),
#     `Sensor` = c(ctd.spec,puck.spec,oxy.spec,par.spec,dmon.spec),
#     `State to Sample` = c(rep("See Table 2",4),state.to.sample.7),
#     #`State to Sample` = c(rep(state.to.sample.7,3),state.to.sample.15),
#     `Depth to Sample` = c(rep(sample.depth,5)),
#     `Serial Number` = c("10049","8869","1132","50423","68")
#     ) %>%
#   knitr::kable(format="latex",align="l",booktabs=TRUE,linesep="") %>%
#   kableExtra::kable_styling(position="left",latex_options="striped",stripe_color="gray!15") %>%
#   column_spec(1,width="1in") %>%
#   column_spec(2,width="2in") %>%
#   column_spec(3,width="1in") %>%
#   column_spec(4,width="1in") %>%
#   column_spec(5,width="1in")
```

## Introduction

The Ecosystem Science Division at NOAA Fisheries’ Southwest Fisheries Science Center monitors the living marine resources within the Southern Ocean and the California Current in order to satisfy the requirements of several legislative mandates to support conservation and management decision-making. To achieve this goal, we use autonomous underwater buoyancy-driven gliders with integrated sensors for measuring ocean conditions, plankton densities, and marine mammal distributions.

Depending on the specific deployment objective, Slocum gliders are equipped with a suite of sensors. We obtain acoustic estimates of zooplankton density (primarily Antarctic krill in the Southern Ocean) using one of two different echosounders: an Acoustic Zooplankton Fish Profiler with discrete frequencies at 67.5 and 125 kHz (AZFP, ASL, Inc) and a mini-Signature 100 wideband echosounder with continuous frequencies between 70 and 120 kHz (Nortek). We also collect ancillary oceanographic data (temperature, salinity, dissolved oxygen, chlorophyll, colored dissolved organic matter, backscatter, and photosynthetically active radiation) to characterize the marine environment. Additional sensors may include passive acoustic monitors for marine mammal detection (“Wispr”, Embedded Ocean Systems; digital acoustic monitoring “DMON”, Woods Hole Oceanographic Institution), “glidercams” for verifying acoustic targets (Williamson and Associates, Inc.) and shadowgraph cameras for obtaining imagery of the plankton community (Williamson and Associates, Inc.). Imagery is used to train artificial intelligence (AI) models to automate plankton identification.

Hefring OceanScout gliders are equipped with conductivity-temperature-depth sensors (CTDs; RBR, Inc.) and “Wispr” passive acoustic monitors.

## Pre-Deployment Preparation and Testing

Prior to deployment, the ESD has a standard protocol for preparing and testing gliders to minimize or eliminate issues that may occur due to human error during deployment:

*Slocum gliders*

1.  Gliders are properly ballasted (i.e., weighted) so that the density of the glider matches the density of the water in which it will be deployed. Weight and flotation configurations are documented
2.  The junctions between glider sections are thoroughly cleaned, old o-rings are discarded, new o-rings are inspected for damage that may compromise their ability to form a water-tight seal, new o-rings are properly lubricated, and the glider is sealed together. All cable connections are photographed to document the “final seal” and ensure the glider was reassembled properly
3.  A “Functional Checkout” is performed to ensure and document that all glider systems and science sensors are functioning properly. During the Functional Checkout, we verify the battery type installed in the glider (lithium primary or lithium rechargeable) and that the appropriate battery duration (total coulomb amp hours) is active in the glider’s autoexec.mi file
4.  Two test missions are performed in the SWFSC test tank (20 m x 10 m x 10 m) to ensure the glider is performing as expected
5.  Once per year, glider compasses are calibrated (the compass was not calibrated prior to this deployment)
6.  Biofouling prevention measures are applied as necessary
7.  Glider flight and science sampling files are prepared according to mission objectives. These objectives are identified by the Principal Investigators for each deployment. Files are uploaded to the Teledyne Webb Research Slocum Fleet Mission Control (SFMC) web interface and sent to the glider just prior to deployment over the Iridium connection
8.  When gliders are shipped to their deployment location, ESD glider technicians perform a second Functional Checkout to ensure the gliders function properly after transit

*OceanScout gliders*

1.  Gliders are configured with the standard ballast for salt water (8 weights in the nose)
2.  O-rings are replaced as needed
3.  Motors (variable buoyancy engine, pitch, and roll) are calibrated, either through the glider’s wifi connection prior to deployment, or by using platform.test through the command line interface
4.  Routes, diving parameters, and sensor sampling strategies are established through the Hefring cloud interface and deployed to the glider

## Deployment-specific testing

This deployment was an initial test of a new shadowgraph model with slightly different mounting configurations. We performed test dives in the SWFSC test tank with the older model’s mounting bracket and the glider dove and climbed as expected. No further deployment-specific testing occurred.

## Deployment

`{r} glider_name` was deployed on `{r} deployment.length[1]`, approximately 9 km west of Oceanside Harbor in Oceanside, CA (`{r} round(nc.eng.lat[1],digits=2)`^o^N, `{r} round(nc.eng.lon[1],digits=2)`^o^W) from the ESD Zodiac R/V *Ernest II*. This glider was deployed with the sensor configuration listed above, and with `{r} battery_type` batteries (coulomb amp hour total = `{r} battery_capacity`).

We began this deployment using the autoballast feature to maximize oil pump efficiency. Initial dives to 20 and 50 m were steep (\~59 degrees) and climbs were shallow (\~16 degrees), but they evened out with deeper dives as autoballast adjusted. Autoballast converged successfully and remained converged for the duration of the deployment.

Because digifins are sensitive to slight movements and produce erroneous oddities that can lead to mission aborts, the following commands were issued to the glider prior to sequencing our standard mission (1k_n.mi):

`put u_digifin_hide_oddities_at_surface 1`

`put u_digifin_mask_movement_warning_at_surface 1`

```{r}
#| echo: false
#| message: false

#for battery usage. Must fill in proper coulomb amphr total for ca.percent.used
nc.eng.cat <- ncvar_get(nc.eng,"total_amphr")
nc.eng.cat <- as.vector(nc.eng.cat)
ca.used <- max(nc.eng.cat)-min(nc.eng.cat)

ca.percent.used <- (ca.used / as.numeric(battery_capacity)) * 100
```

This glider performed well for the entire `{r} days.deployed`-day mission. The glider used `{r} round(ca.used,digits=2)` amp hours over `{r} days.deployed` days, or `{r} round(ca.percent.used,digits=2)`% of its battery capacity.

All science sensors maintained consistent sampling for the duration of the deployment. The shadowgraph was programmed to only collect images up to 300 meters, however the glider never dove deeper than `{r} round(max.depth,digits=1)` meters.

```{r}
# #| label: tbl-sampling
# #| tbl-cap: "Sampling strategies for sensors to determine whether individual sensors interfere with passive acoustic data. Sensors were systematically turned off and on throughout the deployment to isolate interence. Time is in UTC."
# #| echo: false
# #| message: false
# 
# UNCOMMENT FOR RODEO SPECIFIC SAMPLING TABLE
# 
# #read each line of the list as a separate character string
# # all.ma <- lapply(file.path(local.path, listma), readLines) #gcp files
# # all.ma<-lapply(here(base.path,deployment,"archive-sfmc", listma),readLines) #local files
# 
# #unlist the list so that individual strings can be selected
# all.ma<-unlist(all.ma)
# 
# #get rid of leading white spaces by left-justifying text (only an issue for .ma and .dat files)
# all.ma<-str_trim(all.ma,"left")
# 
# #select strings that start with "b_arg:" (ma files),"$" (cfg files), a capital letter (ini files), or that don't start with "#" (proglets)
# ma.args<-all.ma[grep("^b_arg:*",all.ma)]
# 
# #get rid of white space between b_arg and #
# ma.args<-str_squish(ma.args)
# 
# #split character string at "#" (for .ma files only)
# ma.args<-str_split(ma.args,"#")
# 
# #unlist ma.args to be able to get rid of everything after the "#"
# ma.args<-unlist(ma.args)
# 
# #get rid of everything behind the "#"
# ma.args<-ma.args[c(TRUE,FALSE)]
# 
# #create a dataframe so that a column with date/time stamp can be added
# ma.args<-as.data.frame(ma.args)
# 
# #split the original file names to isolate the date/time string
# x1<-str_split(listma,"_")
# 
# #unlist the list so that individual strings can be selected
# x1<-unlist(x1)
# 
# #just select the date/time strings
# x1<-x1[c(TRUE,FALSE)]
# 
# #create objects that repeat the date/time string for the number of arguments per file
# a1<-nrow(ma.args)/length(listma)
# x1<-rep(x1,each=a1)
# 
# #create new columns with date and time for each data frame
# ma.args<-cbind(ma.args,x1)
# 
# #create new data frame with all args
# new.col.names<-c("arg","date.time")
# colnames(ma.args)<-new.col.names
# 
# #split "args" column into multiple columns for ma files
# ma.args<-ma.args %>%
#   separate(arg,c("delete","parameter","value")," ")
# 
# #delete column "b_arg" (not necessary)
# ma.args<-ma.args[-1]
# 
# #create a sensor column for ma.args
# ma.args<-ma.args %>%
#   select(parameter:value,date.time) %>%
#   mutate(
#     Sensor = case_when(
#       parameter == "sensor_type(enum)" & value == 1 ~ "CTD",
#       parameter == "sensor_type(enum)" & value == 48 ~ "ECO Puck",
#       parameter == "sensor_type(enum)" & value == 54 ~ "Optode",
#       parameter == "sensor_type(enum)" & value == 56 ~ "PAR"
#     )
#   )
# 
# #create an "on/off" column"
# ma.args<-ma.args %>%
#   select(parameter:value,date.time,Sensor) %>%
#   mutate(
#     On.Off = case_when(
#       parameter == "state_to_sample(enum)" & value == 1 ~ "On",
#       parameter == "state_to_sample(enum)" & value == 0 ~ "Off"
#     )
#   )
# 
# # shift on.off up by 1
# ma.args$On.Off<-lead(ma.args$On.Off,n=1)
# 
# #get rid of NAs
# ma.args<-na.omit(ma.args)
# 
# #split date and time into two columns
# ma.args<-ma.args %>% separate_wider_delim(date.time,"T",names=c("Date","Time"))
# ma.args<-as.data.frame(ma.args)
# 
# #get rid of first two columns
# ma.args<-ma.args[,-c(1:2)]
# 
# #format date
# ma.args$Date<-strptime(ma.args$Date,format="%Y%m%d",tz="UTC")
# 
# #reformat time
# ma.args$Time<-strptime(ma.args$Time,format="%H%M%S")
# ma.args$Time<-strftime(ma.args$Time,"%H:%M:%S")
# 
# knitr::kable(ma.args,format="latex",booktabs=TRUE,longtable=TRUE,linesep="") %>%
#   kableExtra::kable_styling(position="left",latex_options="striped",stripe_color="gray!15")
```

```{r}
# #| label: tbl-wispr
# #| tbl-cap: "Sensor settings for the 'wispr' passive acoustic monitor to determine optimal settings. Settings were changed approximately every 12 hours for the last three days of the deployment. Time is in UTC."
# #| echo: false
# #| message: false
# 
# UNCOMMENT FOR RODEO SPECIFIC SAMPLING TABLE
# 
# #read each line of the list as a separate character string
# pa.cfg<-lapply(here(base.path,deployment,"archive-sfmc",listpacfg),read_lines)
# 
# #unlist the list so that individual strings can be selected
# pa.cfg<-unlist(pa.cfg)
# 
# #select strings that start with "b_arg:" (ma files),"$" (cfg files), a capital letter (ini files), or that don't start with "#" (proglets)
# pa.cfg.args<-pa.cfg[grep("^\\$ADC,.*",pa.cfg)] 
# 
# #create a dataframe so that a column with date/time stamp can be added
# pa.cfg.args<-as.data.frame(pa.cfg.args)
# 
# #split the original file names to isolate the date/time string
# x2<-str_split(listpacfg,"_")
# 
# #unlist the list so that individual strings can be selected
# x2<-unlist(x2)
# 
# #just select the date/time strings
# x2<-x2[c(TRUE,FALSE)]
# 
# #create new columns with date and time for each data frame
# pa.cfg.args<-cbind(pa.cfg.args,x2)
# 
# #create new data frame with all args
# new.col.names<-c("value","date.time")
# colnames(pa.cfg.args)<-new.col.names
# 
# #splitting date and time
# pa.cfg.args<-pa.cfg.args %>% separate_wider_delim(date.time,"T",names=c("Date","Time"))
# pa.cfg.args<-as.data.frame(pa.cfg.args)
# 
# #reorder columns
# pa.cfg.args<-pa.cfg.args[,c(2,3,1)]
# colnames(pa.cfg.args)<-c("Date","Time","Setting")
# 
# #format date
# pa.cfg.args$Date<-strptime(pa.cfg.args$Date,format="%Y%m%d",tz="UTC")
# 
# #reformat time
# pa.cfg.args$Time<-strptime(pa.cfg.args$Time,format="%H%M%S")
# pa.cfg.args$Time<-strftime(pa.cfg.args$Time,"%H:%M:%S")
# 
# #insert backslashes to prevent values from printing as formulas
# pa.cfg.args$Setting<-paste0("\\",pa.cfg.args$Setting)
# 
# knitr::kable(pa.cfg.args,format="latex",booktabs=TRUE,longtable=TRUE,linesep="") %>%
#   kableExtra::kable_styling(position="left",latex_options="striped",stripe_color="gray!15")
```

## Post-deployment actions

Once the glider was back in the laboratory, pilots inspected and cleaned the glider. All data were offloaded from the science and flight cards, and photos were offloaded from the camera. All files were uploaded to NOAA's Google Cloud Platform ([see Data Management structure and access to data here](https://swfsc.github.io/glider-lab-manual/content/data-management.html#gcs-buckets)). Data will be quality checked and analyzed at a later date.

Pilots checked the shadowgraph photo quality for sharpness, exposure, and artifacts, to assess if settings needed to be adjusted for the next deployment. Based on the image brightness, pilots decided to increase the red flash brightness to 100% for the next deployment, while keeping other camera settings the same.

## Figures

Plots below are generated from raw data which has not yet been quality-checked.

```{r}
if (processor == "GCP") {
  plot_dir <- file.path(base.path, deployment_year, deployment, "plots", "delayed", "spatialGrids-sci")
} else {
  # locally saved plots
  plot_dir <- here(base.path, deployment, "plots", "delayed", "spatialGrids-sci")
}

# Define the individual plot paths
bs   <- file.path(plot_dir, paste0(deployment, "_backscatter_700_spatialGrids.png"))
chl  <- file.path(plot_dir, paste0(deployment, "_chlorophyll_spatialGrids.png"))
den  <- file.path(plot_dir, paste0(deployment, "_density_spatialGrids.png"))
oxy  <- file.path(plot_dir, paste0(deployment, "_oxygen_concentration_spatialGrids.png"))
sal  <- file.path(plot_dir, paste0(deployment, "_salinity_spatialGrids.png"))
temp <- file.path(plot_dir, paste0(deployment, "_temperature_spatialGrids.png"))
```

```{r}
#| label: render-plots
#| echo: false
#| results: asis

plot_list <- list(
  list(path = bs, name = "Backscatter", alt = "A plot showing points where the glider surfaced by latitude and longitude, with two color-coded plots of backscatter by latitude and depth and by longitude and depth."),
  list(path = chl, name = "Chlorophyll-a", alt = "A plot showing points where the glider surfaced by latitude and longitude, with two color-coded plots of chlorophyll-a by latitude and depth and by longitude and depth."),
  list(path = den, name = "Water Density", alt = "A plot showing points where the glider surfaced by latitude and longitude, with two color-coded plots of water density by latitude and depth and by longitude and depth."),
  list(path = oxy, name = "Oxygen", alt = "A plot showing points where the glider surfaced by latitude and longitude, with two color-coded plots of oxygen concentration by latitude and depth and by longitude and depth."),
  list(path = sal, name = "Salinity", alt = "A plot showing points where the glider surfaced by latitude and longitude, with two color-coded plots of salinity by latitude and depth and by longitude and depth."),
  list(path = temp, name = "Temperature", alt = "A plot showing points where the glider surfaced by latitude and longitude, with two color-coded plots of temperature by latitude and depth and by longitude and depth.")
)

for (p in plot_list) {
  if (file.exists(p$path)) {
    cat(paste0('\n\n![', p$alt, '](', p$path, '){width="100%"}\n\n'))
  } else {
    cat(paste0('\n\n*', p$name, ' plot not currently found.*\n\n'))
  }
}
```
